Binocular integration of chromatic and luminance signals

### 2024-03-09 manuscript.Rmd
Started writing manuscript

---
title: "Binocular integration of chromatic information"
author: |
   | Daniel H. Baker, Kirralise J. Hansford, Federico G. Segala, Rowan J. Huxley,   
   | Joel T. Martin, Lauren E. Welbourne & Alex R. Wade
date: "`r Sys.Date()`"
output:
  bookdown::pdf_document2:
    fig_caption: yes
    toc: no
    keep_tex: yes
  word_document: default
  pdf_document:
    toc: no
  html_document: default
bibliography: references.bib
csl: elife.csl
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

>>>>

---
title: "Binocular integration of chromatic and luminance signals"
author: |
   | Daniel H. Baker, Kirralise J. Hansford, Federico G. Segala, Annie Y. Morsi,
   | Rowan J. Huxley, Joel T. Martin, Maya Rockman & Alex R. Wade
date: "`r Sys.Date()`"
output:
  bookdown::pdf_document2:
    fig_caption: yes
    toc: no
    keep_tex: yes
  word_document: default
  html_document: default
bibliography: references.bib
csl: journal-of-vision.csl
header-includes: 
  \usepackage{float} \floatplacement{figure}{H} 
  \newcommand{\beginsupplement}{\setcounter{table}{0}  \renewcommand{\thetable}{A\arabic{table}} \setcounter{figure}{0} \renewcommand{\thefigure}{A\arabic{figure}}}
---

```{r setup, include=FALSE}

processdata <- 0
# the processdata flag has three settings:
# 0: do no analysis, just produce the pdf of the manuscript
# 1: fit psychometric functions and plot graphs
# 2: do Stan modelling (takes N hours)

nsamples <- 10000    # number of Stan samples for modelling

# reasonably compact code to check which packages are installed, install the missing ones, and activate all
packagelist <- c('knitr','reticulate','osfr','bookdown','rstan','coda','modeest','pals','pracma','doParallel','R.matlab') # list of CRAN packages
missingpackages <- packagelist[!packagelist %in% installed.packages()[,1]]
if (length(missingpackages)>0){install.packages(missingpackages)}
toinstall <- packagelist[which(!packagelist %in% (.packages()))]
invisible(lapply(toinstall,library,character.only=TRUE))

# helper function to make colours transparent
addalpha <- function(col, alpha=1){apply(sapply(col, col2rgb)/255, 2, function(x) rgb(x[1], x[2], x[3], alpha=alpha))}

lms2rgb <- function(phosphors,fundamentals,lms){
  rgbTOlms <- t(as.matrix(fundamentals)) %*% as.matrix(phosphors)
  lmsTOrgb <- solve(rgbTOlms)
  rgb <- lmsTOrgb %*% lms
return(rgb)}

getmodelresp <- function(p,L,R){
  Lresp <- (L^p[3])/(p[4] + L + p[6]*R)
  Rresp <- (R^p[3])/(p[4] + R + p[6]*L)
  bs <- Lresp + Rresp
  resp <- (bs^p[1])/(p[5] + bs^p[2])
return(resp)}

discriminate <- function(p,pedC,cond) {

  if (cond==1){baseline <- getmodelresp(p,pedC,0)}
  if (cond==2){baseline <- getmodelresp(p,pedC,pedC)}
  if (cond==3){baseline <- getmodelresp(p,pedC,pedC)}
  if (cond==4){baseline <- getmodelresp(p,pedC,0)}

  modelresp <- -10
  contrastinc <- 0
  if (baseline>-999){
    while (((modelresp-baseline) < p[7])){
      contrastinc <- contrastinc + 0.1
  if (cond==1){modelresp <- getmodelresp(p,pedC+contrastinc,0)}
  if (cond==2){modelresp <- getmodelresp(p,pedC+contrastinc,pedC+contrastinc)}
  if (cond==3){modelresp <- getmodelresp(p,pedC+contrastinc,pedC)}
  if (cond==4){modelresp <- getmodelresp(p,pedC,contrastinc)}
    }
    while (((modelresp-baseline) > p[7])){
      contrastinc <- contrastinc - 0.001
  if (cond==1){modelresp <- getmodelresp(p,pedC+contrastinc,0)}
  if (cond==2){modelresp <- getmodelresp(p,pedC+contrastinc,pedC+contrastinc)}
  if (cond==3){modelresp <- getmodelresp(p,pedC+contrastinc,pedC)}
  if (cond==4){modelresp <- getmodelresp(p,pedC,contrastinc)}
    }    
    }
  else{contrastinc <- 99}
return(contrastinc) }

use_miniconda('r-reticulate')
# install psignifit in the terminal with:
# sudo pip install https://github.com/wichmann-lab/python-psignifit/zipball/master
ps <- import('psignifit')

# calculate constant to scale the slope parameter to the Gaussian SD
C <- ps$utils$my_norminv(1-0.05,0,1) - ps$utils$my_norminv(0.05,0,1)

if (!dir.exists('local/fits')){dir.create('local/fits')}

knitr::opts_chunk$set(echo = TRUE)

```

```{r plotisoluminance, include=FALSE}

if (processdata>0){

fundamentals <- read.csv('local/fundamentals.csv')
phosphors <- read.csv('local/phosphors.csv')
thetalist <- (0:360)*pi/180

pdf('Figures/isosettings.pdf',width=11,height=6)

isodata <- read.csv('local/isosettings.csv')
isomeans <- colMeans(isodata)

collist <- brewer.set1(9)

par(mfrow=c(1,2))

ticklocs <- seq(-1,1,0.5)
plot(x=NULL,y=NULL,axes=FALSE, ann=FALSE, xlim=c(-1,1), ylim=c(-1,1))  
axis(1, at=ticklocs, tck=0.01, lab=F, lwd=2)     
axis(2, at=ticklocs, tck=0.01, lab=F, lwd=2)
mtext(text = ticklocs, side = 1, at=ticklocs, line=0.3)    
mtext(text = ticklocs, side = 2, at=ticklocs, line=0.2, las=1)
title(xlab="L", col.lab=rgb(0,0,0), line=1.5, cex.lab=1.5)   
title(ylab="M", col.lab=rgb(0,0,0), line=1.5, cex.lab=1.5)
title(main='L-M gratings', col.lab=rgb(0,0,0), line=1.5, cex.lab=1.5)

text(-0.95,0.9,'(a)',cex=2)

lines(c(0,0),c(-1,1),lty=2)
lines(c(-1,1),c(0,0),lty=2)

rgbvals <- matrix(0,nrow=length(thetalist),ncol=3)
for (n in 1:length(thetalist)){
    theta <- thetalist[n]
    stimLMS <- c(cos(theta), sin(theta), 0)
    rgbvals[n,] <- lms2rgb(phosphors,fundamentals,stimLMS)
}

rgbvals <- (rgbvals/(2*max(abs(rgbvals)))) + 0.5
for (n in 1:length(thetalist)){    points(cos(thetalist[n]),sin(thetalist[n]),pch=16,col=rgb(rgbvals[n,1],rgbvals[n,2],rgbvals[n,3]),cex=2)}

arctheta <- seq(60,179,0.5)*pi/180
lines(cos(arctheta)*0.45,sin(arctheta)*0.5,lwd=2);
lines(-cos(arctheta)*0.45,-sin(arctheta)*0.5,lwd=2);

for (s in 1:5){
    a <- isomeans[s]*pi/180
    lines(cos(a)*c(-0.9, 0.9),sin(a)*c(-0.9,0.9),lwd=2,col=collist[s]);
}

legend(-1,-0.25,c('P1','P2','P3','P4','P5'),lwd=2,col=collist[1:5],box.lwd=2,cex=1.2,bg='white')

plot(x=NULL,y=NULL,axes=FALSE, ann=FALSE, xlim=c(-1,1), ylim=c(-1,1))  
axis(1, at=ticklocs, tck=0.01, lab=F, lwd=2)     
axis(2, at=ticklocs, tck=0.01, lab=F, lwd=2)
mtext(text = ticklocs, side = 1, at=ticklocs, line=0.3)    
mtext(text = ticklocs, side = 2, at=ticklocs, line=0.2, las=1)
title(xlab="L+M", col.lab=rgb(0,0,0), line=1.5, cex.lab=1.5)   
title(ylab="S", col.lab=rgb(0,0,0), line=1.5, cex.lab=1.5)
title(main='S-(L+M) gratings', col.lab=rgb(0,0,0), line=1.5, cex.lab=1.5)

text(-0.95,0.9,'(b)',cex=2)

lines(c(0,0),c(-1,1),lty=2)
lines(c(-1,1),c(0,0),lty=2)

rgbvals <- matrix(0,nrow=length(thetalist),ncol=3)
for (n in 1:length(thetalist)){
    theta <- thetalist[n]
    stimLMS <- c(cos(theta), cos(theta), sin(theta))
    rgbvals[n,] <- lms2rgb(phosphors,fundamentals,stimLMS)
}

rgbvals <- (rgbvals/(2*max(abs(rgbvals)))) + 0.5
for (n in 1:length(thetalist)){    points(cos(thetalist[n]),sin(thetalist[n]),pch=16,col=rgb(rgbvals[n,1],rgbvals[n,2],rgbvals[n,3]),cex=2)}

arctheta <- seq(45,135,0.5)*pi/180
lines(cos(arctheta)*0.45,sin(arctheta)*0.5,lwd=2);
lines(-cos(arctheta)*0.45,-sin(arctheta)*0.5,lwd=2);

for (s in 1:5){
    a <- isomeans[s+5]*pi/180
    lines(cos(a)*c(-0.9, 0.9),sin(a)*c(-0.9,0.9),lwd=2,col=collist[s]);
}

dev.off()

}

```

```{r loaddata, include=FALSE}

ex1data <- read.csv('local/Experiment1.csv')
ex2data <- read.csv('local/Experiment2.csv')

if (processdata>0){

sublist <- unique(ex1data$Subject)
explist <- unique(ex1data$Experiment)

options <- NULL
options$sigmoidName <- 'norm'   # choose a cumulative Gauss as the sigmoid
options$expType <- '2AFC'   # choose 2-AFC as the paradigm of the experiment

allthresh <- array(0,dim=c(3,3,4,8))
allslope <- array(0,dim=c(3,3,4,8))
for (subj in 1:3){
  subdata <- subset(ex1data,Subject==sublist[subj])
for (expt in 1:3){
  expdata <- subset(subdata,Experiment==explist[expt])
for (cond in 1:4){
  conddata <- subset(expdata,Condition==(2*cond)|Condition==((2*cond)-1))
  pedlevs <- unique(conddata$PedestalContrast)
for (pedlevel in 1:8){

blockdata <- subset(conddata,PedestalContrast==pedlevs[pedlevel])

if (nrow(blockdata)>0){
targetcontrasts <- sort(unique(blockdata$TargetContrast))
ntrials <- rep(0,length(targetcontrasts))
ncorrect <- rep(0,length(targetcontrasts))
for (lev in 1:length(targetcontrasts)){
  temp <- subset(blockdata,TargetContrast==targetcontrasts[lev])
  ntrials[lev] <- nrow(temp)
  ncorrect[lev] <- sum(temp$IsCorrect)
}

tofit <- as.matrix(data.frame(20*log10(targetcontrasts),ncorrect,ntrials))

res <- ps$psignifit(tofit,options)

allthresh[subj,expt,cond,pedlevel] <- res$Fit[1]
allslope[subj,expt,cond,pedlevel] <- res$Fit[2]/C
}

}}}}

# shift hbin and dich conditions up one
allthresh[,,3,2:8] <- allthresh[,,3,1:7]
allthresh[,,4,2:8] <- allthresh[,,4,1:7]
allslope[,,3,2:8] <- allslope[,,3,1:7]
allslope[,,4,2:8] <- allslope[,,4,1:7]
# duplicate the mon threshold conditions for hbin and dich
allthresh[,,3,1] <- allthresh[,,1,1]
allthresh[,,4,1] <- allthresh[,,1,1]
allslope[,,3,1] <- allslope[,,1,1]
allslope[,,4,1] <- allslope[,,1,1]

meanthresh <- apply(allthresh,2:4,mean)
meanslope <- apply(allslope,2:4,mean)

SEthresh <- apply(allthresh,2:4,sd)/sqrt(3)
SEslope <- apply(allslope,2:4,sd)/sqrt(3)

save(file='local/thresholddata.RData',list=c('allthresh','allslope','meanthresh','meanslope','SEthresh','SEslope'))

pedlist <- c('DET','AC','RG','BY')
tarlist <- c('AC','RG','BY')
sublist <- unique(ex2data$Subject)

ntrials <- array(0,dim=c(length(sublist),12,10))
ncorrect <- ntrials
contrastsC <- ntrials
counter <- 0
for (ped in 1:length(pedlist)){
  for (tar in 1:length(tarlist)){
    counter <- counter + 1
    for (subj in 1:length(sublist)){
    thiscond <- subset(ex2data,Subject==sublist[subj] & TargetType==tarlist[tar] & PedestalType==pedlist[ped])
    targetcontrasts <- sort(as.numeric(unique(thiscond$TargetContrast)))
    for (tlevel in 1:length(targetcontrasts)){
      thislevel <- subset(thiscond, TargetContrast==targetcontrasts[tlevel])
      ntrials[subj,counter,tlevel] <- nrow(thislevel)
      ncorrect[subj,counter,tlevel] <- sum(thislevel$IsCorrect)
      contrastsC[subj,counter,tlevel] <- targetcontrasts[tlevel]
    }
  }
}
}

MCSpropcorr <- 100*ncorrect/ntrials
contrastsdB <- round(20*log10(contrastsC))

MCSthresh <- matrix(0,nrow=length(sublist),ncol=12)
MCSslope <- MCSthresh
for (subj in 1:length(sublist)){
for (cond in 1:12){
  datatofit <- data.frame(contrastsdB[subj,cond,],ncorrect[subj,cond,],ntrials[subj,cond,])
  colnames(datatofit) <- c('CdB','Ncorrect','Ntrials')

  res <- ps$psignifit(as.matrix(datatofit),options)

  MCSthresh[subj,cond] <- res$Fit[1]
  MCSslope[subj,cond] <- res$Fit[2]/C

}}

thdelev <- array(0,dim=c(3,3,3))
slopematrix <- array(0,dim=c(3,3,3))

for (subj in 1:length(sublist)){
thdelev[subj,1,] <- MCSthresh[subj,c(4,7,10)] - MCSthresh[subj,1]
thdelev[subj,2,] <- MCSthresh[subj,c(5,8,11)] - MCSthresh[subj,2]
thdelev[subj,3,] <- MCSthresh[subj,c(6,9,12)] - MCSthresh[subj,3]

slopematrix[subj,1,] <- 10.3/MCSslope[subj,c(4,7,10)]
slopematrix[subj,2,] <- 10.3/MCSslope[subj,c(5,8,11)]
slopematrix[subj,3,] <- 10.3/MCSslope[subj,c(6,9,12)]
}
detslopes <- 10.3/MCSslope[,1:3]

save(file='local/MCSdata.RData',list=c('MCSthresh','MCSslope','thdelev','slopematrix','detslopes','MCSpropcorr','contrastsdB'))

}

```

```{r domodellingsimplex, include=FALSE}

load('local/thresholddata.RData')

getmodelresp <- function(p,L,R){
  Lresp <- (L^p[3])/(p[4] + L + p[6]*R)
  Rresp <- (R^p[3])/(p[4] + R + p[6]*L)
  bs <- Lresp + Rresp
  resp <- (bs^p[1])/(p[5] + bs^p[2])
return(resp)}

discriminate <- function(p,pedC,cond) {

  if (cond==1){baseline <- getmodelresp(p,pedC,0)}
  if (cond==2){baseline <- getmodelresp(p,pedC,pedC)}
  if (cond==3){baseline <- getmodelresp(p,pedC,pedC)}
  if (cond==4){baseline <- getmodelresp(p,pedC,0)}

  modelresp <- -10
  contrastinc <- 0
  if (baseline>-999){

    while (((modelresp-baseline) < p[7])){
      contrastinc <- contrastinc + 0.1
  if (cond==1){modelresp <- getmodelresp(p,pedC+contrastinc,0)}
  if (cond==2){modelresp <- getmodelresp(p,pedC+contrastinc,pedC+contrastinc)}
  if (cond==3){modelresp <- getmodelresp(p,pedC+contrastinc,pedC)}
  if (cond==4){modelresp <- getmodelresp(p,pedC,contrastinc)}
    }
    while (((modelresp-baseline) > p[7])){
      contrastinc <- contrastinc - 0.001
  if (cond==1){modelresp <- getmodelresp(p,pedC+contrastinc,0)}
  if (cond==2){modelresp <- getmodelresp(p,pedC+contrastinc,pedC+contrastinc)}
  if (cond==3){modelresp <- getmodelresp(p,pedC+contrastinc,pedC)}
  if (cond==4){modelresp <- getmodelresp(p,pedC,contrastinc)}
    }    
    }
  else{contrastinc <- 99}
return(contrastinc) }

errorfit <- function(p){

p <- 10^p
p[2] <- p[2] + 1
p[2] <- min(p[2],10)
p[1] <- 1 + p[1] + p[2]
p[3] <- 1 + p[3]

pedlevelsC <- 10^(pedlist/20)
pedlevelsC[1] <- 0
allpred <- array(0,dim=c(4,length(pedlist)))

  for (cond in 1:4){
    for (pedlev in 1:length(pedlist)){
      allpred[cond,pedlev] <- discriminate(p,pedlevelsC[pedlev],cond)
    }
  }

allpreddB <- 20*log10(allpred)
rms <- sqrt(mean((allpreddB - datatofit)^2))
if (is.na(rms)){rms <- 999}
return(rms)}

if (processdata==2){

ntotalsimplexfits <- 100

cluster <- makeCluster(8)
registerDoParallel(cluster)

pedlist <<- seq(-12,30,6)
allp <- matrix(0,nrow=3,ncol=7)
allparams <- array(0,dim=c(3,ntotalsimplexfits,8))
allrms <- NULL
 for (cond in 1:3){
   print(cond)
   datatofit <<- meanthresh[cond,,]

   nfits <- ntotalsimplexfits
   for (n in 1:ntotalsimplexfits){
     if (file.exists(paste0('local/fits/E',cond,'f',n,'.RData'))){nfits <- nfits-1}
   }

   if (nfits > 0){
   foreach(i = 1:nfits, .combine=rbind) %dopar% {
   sout <- pracma::nelder_mead(errorfit,rnorm(7,sd=0.1)+log10(c(0.5,5.5,0.3,1,0.01,1,0.2)))
   allout <- c(errorfit(sout$xmin),sout$xmin)

   hassaved <- 0
   filecount <- 0
   while (!hassaved){
     filecount <- filecount + 1
     if (!file.exists(paste0('local/fits/E',cond,'f',filecount,'.RData'))){
       save(file=paste0('local/fits/E',cond,'f',filecount,'.RData'),list='allout')
       hassaved <- 1}
   }
   allout <- c(errorfit(sout$xmin),sout$xmin)
   }
   }

   finalout <- matrix(0,nrow=ntotalsimplexfits,ncol=8)
   for (n in 1:ntotalsimplexfits){
     if (file.exists(paste0('local/fits/E',cond,'f',n,'.RData'))){
       load(file=paste0('local/fits/E',cond,'f',n,'.RData'))
       finalout[n,] <- allout
     }
   }
   i <- which(finalout[,1]==min(finalout[,1]))
   p <- 10^finalout[i[1],2:8]
   allrms[cond] <- finalout[i[1],1]
   p[2] <- p[2] + 1
   p[2] <- min(p[2],10)
   p[1] <- 1 + p[1] + p[2]
   p[3] <- 1 + p[3]
   allp[cond,] <- p
   pedlist <<- seq(-6,36,6)

   allparams[cond,,] <- finalout
 }

save(file='local/simplexfits.RData',list=c('allp','allrms','allparams','ntotalsimplexfits'))

stopCluster(cluster)

}

```

```{r plotmodellingsimplex, include=FALSE}

if (processdata>0){

load('local/simplexfits.RData')

pedlist <- -12:39
pedlevelsC <- 10^(pedlist/20)
pedlevelsC[1] <- 0
allpred <- array(0,dim=c(3,4,length(pedlist)))

for (expt in 1:3){
  for (cond in 1:4){
    for (pedlev in 1:length(pedlist)){
      allpred[expt,cond,pedlev] <- discriminate(allp[expt,],pedlevelsC[pedlev],cond)
    }
  }
}
allpreddB <- 20*log10(allpred)

pdf('Figures/dipperssimplex.pdf',width=12,height=5)

collist <- c('red','blue','orange','darkgreen')

ticklocsx <- seq(-12,36,6)
ticklabelsx <- c(0,0.5,1,2,4,8,16,32,64)
ticklocsy <- seq(-12,36,6)
ticklabelsy <- c(0.25,0.5,1,2,4,8,16,32,64)

par(pty="s",mfcol=c(1,3))  # make axis square
plot(x=NULL,y=NULL,axes=FALSE, ann=FALSE, xlim=c(-12,36), ylim=c(-12,39))   
axis(1, at=ticklocsx, tck=0.01, lab=F, lwd=2)     # plot tick marks (no labels)
axis(2, at=ticklocsy, tck=0.01, lab=F, lwd=2)
mtext(text = ticklabelsx, side = 1, at=ticklocsx, line=0.3)     
mtext(text = ticklabelsy, side = 2, at=ticklocsy, line=0.2, las=1)
title(xlab="Pedestal contrast (%)", col.lab=rgb(0,0,0), line=1.5, cex.lab=1.5)   
title(ylab="Threshold (%)", col.lab=rgb(0,0,0), line=1.5, cex.lab=1.5)
title(main='Achromatic')

pedcontrasts <- seq(-12,30,6)

for (cond in 1:4){
  arrows(pedcontrasts,meanthresh[1,cond,],pedcontrasts,meanthresh[1,cond,]+SEthresh[1,cond,],col=collist[cond],angle=90,length=0.02,lwd=2)
 arrows(pedcontrasts,meanthresh[1,cond,],pedcontrasts,meanthresh[1,cond,]-SEthresh[1,cond,],col=collist[cond],angle=90,length=0.02,lwd=2)  
  # lines(pedcontrasts,meanthresh[1,cond,],col=collist[cond],lwd=2)
  lines(pedlist,allpreddB[1,cond,],col=collist[cond],lwd=2)
  points(pedcontrasts,meanthresh[1,cond,],pch=16,col=collist[cond],cex=1.5)
}
text(24,-9,paste0('RMS = ',round(allrms[1],digits=2),'dB'),cex=2)
legend(-12,36,c('Monocular','Binocular','Half-bin','Dichoptic'),pch=16,col=collist,box.lwd=2,cex=1.5)

par(pty="s")  # make axis square
plot(x=NULL,y=NULL,axes=FALSE, ann=FALSE, xlim=c(-12,36), ylim=c(-12,39))  
axis(1, at=ticklocsx, tck=0.01, lab=F, lwd=2)     # plot tick marks (no labels)
axis(2, at=ticklocsy, tck=0.01, lab=F, lwd=2)
mtext(text = ticklabelsx, side = 1, at=ticklocsx, line=0.3)    
mtext(text = ticklabelsy, side = 2, at=ticklocsy, line=0.2, las=1)
title(xlab="Pedestal contrast (%)", col.lab=rgb(0,0,0), line=1.5, cex.lab=1.5)   
title(ylab="Threshold (%)", col.lab=rgb(0,0,0), line=1.5, cex.lab=1.5)
title(main='Red/Green')

pedcontrasts <- c(-12,seq(0,36,6))

for (cond in 1:4){
  arrows(pedcontrasts,meanthresh[3,cond,],pedcontrasts,meanthresh[3,cond,]+SEthresh[3,cond,],col=collist[cond],angle=90,length=0.02,lwd=2)
 arrows(pedcontrasts,meanthresh[3,cond,],pedcontrasts,meanthresh[3,cond,]-SEthresh[3,cond,],col=collist[cond],angle=90,length=0.02,lwd=2)  
 # lines(pedcontrasts,meanthresh[3,cond,],col=collist[cond],lwd=2)
  lines(pedlist,allpreddB[3,cond,],col=collist[cond],lwd=2)
  points(pedcontrasts,meanthresh[3,cond,],pch=16,col=collist[cond],cex=1.5)
}
text(24,-9,paste0('RMS = ',round(allrms[2],digits=2),'dB'),cex=2)

par(pty="s")  # make axis square
plot(x=NULL,y=NULL,axes=FALSE, ann=FALSE, xlim=c(-12,36), ylim=c(-12,39))   
axis(1, at=ticklocsx, tck=0.01, lab=F, lwd=2)     # plot tick marks (no labels)
axis(2, at=ticklocsy, tck=0.01, lab=F, lwd=2)
mtext(text = ticklabelsx, side = 1, at=ticklocsx, line=0.3)     
mtext(text = ticklabelsy, side = 2, at=ticklocsy, line=0.2, las=1)
title(xlab="Pedestal contrast (%)", col.lab=rgb(0,0,0), line=1.5, cex.lab=1.5)
title(ylab="Threshold (%)", col.lab=rgb(0,0,0), line=1.5, cex.lab=1.5)
title(main='Blue/Yellow')

for (cond in 1:4){
  arrows(pedcontrasts,meanthresh[2,cond,],pedcontrasts,meanthresh[2,cond,]+SEthresh[2,cond,],col=collist[cond],angle=90,length=0.02,lwd=2)
 arrows(pedcontrasts,meanthresh[2,cond,],pedcontrasts,meanthresh[2,cond,]-SEthresh[2,cond,],col=collist[cond],angle=90,length=0.02,lwd=2)  
 # lines(pedcontrasts,meanthresh[2,cond,],col=collist[cond],lwd=2)
  lines(pedlist,allpreddB[2,cond,],col=collist[cond],lwd=2)
  points(pedcontrasts,meanthresh[2,cond,],pch=16,col=collist[cond],cex=1.5)
}
text(24,-9,paste0('RMS = ',round(allrms[3],digits=2),'dB'),cex=2)

dev.off()

}

```

```{r domodelling, include=FALSE}

if (processdata==2){

  numSavedSteps <- nsamples    # total number of MCMC steps
  adaptSteps = 200  # Number of steps to "tune" the samplers
  burnInSteps = 200 
  thinSteps <- 1
  nChains = 24
  options(mc.cores=8)

modelString = "
  data {
    int<lower=1> Nsubj ;
    int<lower=1> Ntotal ;
    int correct[Ntotal] ;
    real PedestalContrast[Ntotal] ;
    real TargetContrast[Ntotal] ;
    int<lower=1> c[Ntotal] ;
    int<lower=1> s[Ntotal] ;
  }
  parameters {
    real<lower=0> p[Nsubj] ;
    real<lower=0> q[Nsubj] ;    
    real<lower=0> m[Nsubj] ;
    real<lower=0> S[Nsubj] ;  
    real<lower=0> Z[Nsubj] ;
    real<lower=0> w[Nsubj] ;
    real<lower=0> k[Nsubj] ;

    real<lower=0> sigma ;
    real<lower=0> pmu ; 
    real<lower=0> qmu ; 
    real<lower=0> mmu ;
    real<lower=0> Smu ;
    real<lower=0> Zmu ; 
    real<lower=0> wmu ; 
    real<lower=0> kmu ;

    real<lower=0> psigma ;
    real<lower=0> qsigma ;
    real<lower=0> msigma ;
    real<lower=0> Ssigma ;
    real<lower=0> Zsigma ;    
    real<lower=0> wsigma ;
    real<lower=0> ksigma ;    
  }
  model {

    pmu ~ normal( 1.4 , 1.4/3 ) ;
    qmu ~ normal( 6.59 , 6.59/3 ) ;
    mmu ~ normal( 1.28 , 1.28/3 ) ;
    Smu ~ normal( 0.985 , 0.985/3 ) ;
    Zmu ~ uniform( 0.001 , 100 ) ;
    wmu ~ normal( 1 , 0.3333 ) ;
    kmu ~ normal( 0.194 , 0.194/2 ) ;
    sigma ~ uniform( 1.0E-3 , 1.0E+3 ) ;
    psigma ~ uniform( 1.0E-3 , 1.0E+3 ) ;
    qsigma ~ uniform( 1.0E-3 , 1.0E+3 ) ;
    msigma ~ uniform( 1.0E-3 , 1.0E+3 ) ;
    Ssigma ~ uniform( 1.0E-3 , 1.0E+3 ) ;
    Zsigma ~ uniform( 1.0E-3 , 1.0E+3 ) ;    
    wsigma ~ uniform( 1.0E-3 , 1.0E+3 ) ;
    ksigma ~ uniform( 1.0E-3 , 1.0E+3 ) ;

    p ~ normal( pmu , psigma ) ; // vectorized
    q ~ normal( qmu , qsigma ) ; // vectorized
    m ~ normal( mmu , msigma ) ; // vectorized
    S ~ normal( Smu , Ssigma ) ; // vectorized
    Z ~ normal( Zmu , Zsigma ) ; // vectorized
    w ~ normal( wmu , wsigma ) ; // vectorized
    k ~ normal( kmu , ksigma ) ; // vectorized

    for ( i in 1:Ntotal ) {

    real Lt ;
    real Rt ;
    real Ln ;
    real Rn ;
    real targresp ;
    real nullresp ;

    if (c[i]==1) { // monocular condition
    Lt = pow(PedestalContrast[i]+TargetContrast[i],m[s[i]]) / (S[s[i]] + (PedestalContrast[i]+TargetContrast[i])) ;
    Rt = 0 ;
    Ln = pow(PedestalContrast[i],m[s[i]]) / (S[s[i]] + PedestalContrast[i]) ;
    Rn = 0 ; 
    }

    if (c[i]==2) { // binocular condition 
      Lt = pow(PedestalContrast[i]+TargetContrast[i],m[s[i]]) / (S[s[i]] + (PedestalContrast[i]+TargetContrast[i]) + w[s[i]]*(PedestalContrast[i]+TargetContrast[i])) ;
      Rt = pow(PedestalContrast[i]+TargetContrast[i],m[s[i]]) / (S[s[i]] + (PedestalContrast[i]+TargetContrast[i]) + w[s[i]]*(PedestalContrast[i]+TargetContrast[i])) ;
      Ln = pow(PedestalContrast[i],m[s[i]]) / (S[s[i]] + PedestalContrast[i] + w[s[i]]*PedestalContrast[i]) ;
      Rn = pow(PedestalContrast[i],m[s[i]]) / (S[s[i]] + PedestalContrast[i] + w[s[i]]*PedestalContrast[i]) ;
    }

    if (c[i]==3) { // half-binocular condition 
      Lt = pow(PedestalContrast[i]+TargetContrast[i],m[s[i]]) / (S[s[i]] + (PedestalContrast[i]+TargetContrast[i]) + w[s[i]]*PedestalContrast[i]) ;
      Rt = pow(PedestalContrast[i],m[s[i]]) / (S[s[i]] + PedestalContrast[i] + w[s[i]]*(PedestalContrast[i]+TargetContrast[i])) ;
      Ln = pow(PedestalContrast[i],m[s[i]]) / (S[s[i]] + PedestalContrast[i] + w[s[i]]*PedestalContrast[i]) ;
      Rn = pow(PedestalContrast[i],m[s[i]]) / (S[s[i]] + PedestalContrast[i] + w[s[i]]*PedestalContrast[i]) ;  
    }

    if (c[i]==4) { // dichoptic condition 
      Lt = pow(TargetContrast[i],m[s[i]]) / (S[s[i]] + TargetContrast[i] + w[s[i]]*PedestalContrast[i]) ;
      Rt = pow(PedestalContrast[i],m[s[i]]) / (S[s[i]] + PedestalContrast[i] + w[s[i]]*TargetContrast[i]) ;
      Ln = 0 ;
      Rn = pow(PedestalContrast[i],m[s[i]]) / (S[s[i]] + PedestalContrast[i]) ;      
    }

    targresp = pow((Lt + Rt), (p[s[i]]+q[s[i]])) / (Z[s[i]] + pow((Lt + Rt), q[s[i]])) ;
    nullresp = pow((Ln + Rn), (p[s[i]]+q[s[i]])) / (Z[s[i]] + pow((Ln + Rn), q[s[i]])) ;     
    correct[i] ~ bernoulli((1 + erf((targresp - nullresp)/k[s[i]]))/2) ;

    }
  }  
  " # close quote for modelString

  # Translate to C++ and compile to DSO:
    dipmodel <- stan_model(model_code=modelString)

  ex1data <- read.csv('local/Experiment1.csv')
  explist <- unique(ex1data$Experiment)

for (mod in 1:1){
  expdata <- subset(ex1data,Experiment==explist[mod])
  expdata$Participant <- as.numeric(gsub('\\D','',expdata$Subject))
  expdata$Condition <- ceiling(expdata$Condition/2)

  dataList = list(correct = expdata$IsCorrect, PedestalContrast = expdata$PedestalContrast, TargetContrast = expdata$TargetContrast, c = expdata$Condition, s = expdata$Participant, Nsubj = max(expdata$Participant), Ntotal = nrow(expdata))

  # Get MC sample of posterior:
    stanFit <- sampling(object=dipmodel, 
                        data = dataList, 
                        chains = nChains,
                        iter = (ceiling(numSavedSteps/nChains)*thinSteps
                                +burnInSteps), 
                        warmup = burnInSteps, 
                        thin = thinSteps,
                        cores = getOption("mc.cores", 1L))

    # convert stan format to coda format:jkn9
    mcmcCoda2 = mcmc.list(lapply( 1:ncol(stanFit), function(x){mcmc(as.array(stanFit)[,x,])}))

    save(file=paste0('local/E1mod',mod,'.RData'),list=c('stanFit','mcmcCoda2'))

}

}

```

```{r plotdata, include=FALSE}

load('local/thresholddata.RData')

if (processdata>0){

  paramorder <- c('p','q','m','S','Z','w','k')
  allsamples <- array(0,dim=c(3,7,10008))
  for (expt in 1:3){
    load(file=paste0('local/E1mod',expt,'.RData'))
    p <- NULL
    q <- NULL
    m <- NULL
    S <- NULL
    Z <- NULL
    w <- NULL
    k <- NULL
    for (i in 1:length(mcmcCoda2)){
      p <- c(p,mcmcCoda2[[i]][,23])  
      q <- c(q,mcmcCoda2[[i]][,24])  
      m <- c(m,mcmcCoda2[[i]][,25])  
      S <- c(S,mcmcCoda2[[i]][,26])  
      Z <- c(Z,mcmcCoda2[[i]][,27])  
      w <- c(w,mcmcCoda2[[i]][,28])  
      k <- c(k,mcmcCoda2[[i]][,29])  
    }        
    p <- p + q
allsamples[expt,1,] <- p
allsamples[expt,2,] <- q
allsamples[expt,3,] <- m
allsamples[expt,4,] <- S
allsamples[expt,5,] <- Z
allsamples[expt,6,] <- w
allsamples[expt,7,] <- k

  }
# meanparams <- 10^apply(log10(allsamples),1:2,mlv,method='meanshift')
meanparams <- 10^apply(log10(allsamples),1:2,mean)
# meanparams[2:3,5] <- meanparams[2:3,5]*100

pdf('Figures/modelparams.pdf',width=12,height=8)

collist <- c('black','cornflowerblue','darkred')

par(pty="s",mfcol=c(2,4))

ticklocsx <- seq(-18,18,6)
ticklabelsx <- c('1/8','1/4','1/2',1,2,4,8)
ticklocsy <- seq(0,1,0.2)
ticklabelsy <- ticklocsy

for (param in 1:7){

  plot(x=NULL,y=NULL,axes=FALSE, ann=FALSE, xlim=c(-18,18), ylim=c(0,1))  
  axis(1, at=ticklocsx, tck=0.01, lab=F, lwd=2)     
axis(2, at=ticklocsy, tck=0.01, lab=F, lwd=2)
mtext(text = ticklabelsx, side = 1, at=ticklocsx, line=0.3)    
mtext(text = ticklabelsy, side = 2, at=ticklocsy, line=0.2, las=1)
title(xlab=paramorder[param], col.lab=rgb(0,0,0), line=1.5, cex.lab=1.5)

  for (expt in 1:3){

    svals <- 20*log10(allsamples[expt,param,])
    a <- density(svals)
    a$y <- a$y/max(a$y)
    lines(a$x,a$y,col=collist[expt])

    lines(20*log10(meanparams[expt,c(param,param)]),c(0,1),col=collist[expt])
  }

}

dev.off()
